\huge{Comment on `The Limits of Cosmic Shear'}

### 2016-12-05 cfhtlens.tex
minor update

The Canada-France-Hawaii Telescope Lensing Survey (CFHTLenS) represented a major step forward for the field of weak gravitational lensing, in terms of improved accuracy in data reduction \citep{erben/etal:2013}, the implementation of Gaussianised matched multi-band photometry \citep{hildebrandt/etal:2012}, cross-correlation clustering analysis between photometric redshift slices to verify tomographic redshift distributions \citep{benjamin/etal:2013}, accurate calibrated shape measurements \citep{miller/etal:2013} and a full suite of informative systematic tests to select a clean data set \citep{heymans/etal:2012}.    Since the public release of this survey in 2013, the community has continued to scrutinise and advance our understanding of CFHTLenS by identifying a number of areas where the analysis could improve:
\begin{itemize}
\item{\citet{choi/etal:2016} identified biases in the tomographic photometric redshift distributions using a more effective clustering analysis, in comparison to \citet{benjamin/etal:2013}, by incorporating newly overlapping spectroscopy from the Sloan Digital Sky Survey.  The CFHTLenS tomographic cosmological analysis was then revisited by \citet{joudaki/etal:2016} in order to include a full redshift error analysis based on the results from \citet{choi/etal:2016}, discussed further in section~\ref{sec:photoz}.}
\item{\citet{asgari/etal:2016} used the stringent `COSEBI' statistic to identify significant non-lensing `B-mode' distortions when the CFHTLenS data was split into tomographic slices.}
\item{\citet{kuijken/etal:2015} showed that the CFHTLenS shear calibration corrections derived in \citet{miller/etal:2013} were underestimated as a result of an imperfect match between the galaxy populations in the data and image simulations.}
\item{\citet{fenechconti/etal:2016} demonstrated that the CFHTLenS data would have been subject to a weight bias that favours galaxies that are more intrinsically oriented with the point-spread function.}
\item{\citet{joudaki/etal:2016} updated the CFHTLenS covariance matrices using larger-box numerical simulations that were less subject to the lack power on large scales.}
\end{itemize}
All these advances in our understanding were incorporated and accounted for in the recent KiDS cosmic shear analysis \citep{hildebrandt/etal:2016} which reports at $2.3 \sigma$ tension with Planck.  Efforts are now underway to fully re-analyse CFHTLenS using the advanced KiDS analysis pipeline with revised shape measurements and calibrations for the shear and photometric redshifts.

Given the known shortcomings with the original CFHTLenS results, which will impact in different ways on the cosmological conclusions that one can draw from the survey, it is surprising that \citet{kitching/etal:2016} chose to focus their study on the cosmic shear analysis of \citet{kilbinger/etal:2013}.  This is even more unexpected as, out of all the CFHTLenS cosmic shear results, the non-tomographic analysis of \citet{kilbinger/etal:2013} is the least in tension with Planck \citep[see for example][who show that with $\sigma_8 (\Omega_m/0.3)^{0.5} = 0.738^{+0.055}_{-0.032}$, \citet{kilbinger/etal:2013} is in agreement with Planck at $1.6 \sigma$]{abbott/etal:2016}.

\citet{kitching/etal:2016} choose to only vary $\sigma_8$ in their analysis, fixing all other parameters to Planck constraints.  They find a conditional value of $\sigma_8 = 0.789 \pm 0.015$ for their `standard' analysis (row 1 in Table~\ref{tab:Tl_nu}) of the data from \citet{kilbinger/etal:2013}.    We argue that this artificial reduction in the error by a factor of $\sim 3$ could easily mislead, making it appear that the impact of removing all approximations from the theoretical analysis is more significant for current surveys than it is in reality.

>>>>

The Canada-France-Hawaii Telescope Lensing Survey (CFHTLenS) represented a major step forward for the field of weak gravitational lensing, in terms of improved accuracy in data reduction \citep{erben/etal:2013}, the implementation of Gaussianised matched multi-band photometry \citep{hildebrandt/etal:2012}, cross-correlation clustering analysis between photometric redshift slices to verify tomographic redshift distributions \citep{benjamin/etal:2013}, accurate calibrated shape measurements \citep{miller/etal:2013} and a full suite of informative systematic tests to select a clean data set \citep{heymans/etal:2012}.    Since the public release of this survey in 2013, the community has continued to scrutinise and advance our understanding of CFHTLenS by identifying a number of areas where the analysis could improve:
\begin{itemize}
\item{\citet{choi/etal:2016} identified biases in the tomographic photometric redshift distributions using a more effective clustering analysis, in comparison to \citet{benjamin/etal:2013}, by incorporating newly overlapping spectroscopy from the Sloan Digital Sky Survey.  The CFHTLenS tomographic cosmological analysis was then revisited by \citet{joudaki/etal:2016} in order to include a full redshift error analysis based on the results from \citet{choi/etal:2016}, discussed further in section~\ref{sec:photoz}.}
\item{\citet{asgari/etal:2016} used the stringent `COSEBI' statistic to identify significant non-lensing `B-mode' distortions when the CFHTLenS data was split into tomographic slices.}
\item{\citet{kuijken/etal:2015} showed that the CFHTLenS shear calibration corrections derived in \citet{miller/etal:2013} were underestimated as a result of an imperfect match between the galaxy populations in the data and image simulations.}
\item{\citet{fenechconti/etal:2016} demonstrated that the CFHTLenS data would have been subject to a weight bias that favours galaxies that are more intrinsically oriented with the point-spread function.}
\item{\citet{joudaki/etal:2016} updated the CFHTLenS covariance matrices using larger-box numerical simulations that were less subject to the lack of power on large scales.}
\end{itemize}
All these advances in our understanding were incorporated and accounted for in the recent KiDS cosmic shear analysis \citep{hildebrandt/etal:2016} which reports a $2.3 \sigma$ tension with Planck.  Efforts are now underway to fully re-analyse CFHTLenS using the advanced KiDS analysis pipeline with revised shape measurements and calibrations for the shear and photometric redshifts.

Given the known shortcomings with the original CFHTLenS results, which will impact in different ways on the cosmological conclusions that one can draw from the survey, it is surprising that \citet{kitching/etal:2016} chose to focus their study on the cosmic shear analysis of \citet{kilbinger/etal:2013}.  This is even more unexpected as, out of all the CFHTLenS cosmic shear results, the non-tomographic analysis of \citet{kilbinger/etal:2013} is the least in tension with Planck \citep[see for example][who show that with $\sigma_8 (\Omega_m/0.3)^{0.5} = 0.738^{+0.055}_{-0.032}$, \citet{kilbinger/etal:2013} is in agreement with Planck at $1.6 \sigma$]{abbott/etal:2016}. \shahab{(not sure I entirely agree with this last sentence, please see email)}

\citet{kitching/etal:2016} choose to only vary $\sigma_8$ in their analysis, fixing all other parameters to their Planck values. They find a conditional value of $\sigma_8 = 0.789 \pm 0.015$ for their `standard' analysis (row 1 in Table~\ref{tab:Tl_nu}) of the data from \citet{kilbinger/etal:2013}.    We argue that this artificial reduction in the error by a factor of $\sim 3$ could easily mislead, making it appear that the impact of removing all approximations from the theoretical analysis is more significant for current surveys than it is in reality.

### 2016-12-05 conclusion.tex
minor update

In this comment we have highlighted a number of shortcomings in the analysis presented in \cite{kitching/etal:2016}.  We argue that the flat sky and Limber approximations that can be used in cosmic shear analyses are not able to explain away the tension between current weak lensing and Planck results, in contrast to the message that \cite{kitching/etal:2016} convey.  Our critique should not, however, detract from the important core message of their paper, that future surveys will need to consider these approximations and optimise their statistical analyses accordingly.  For example moving from the standard two-point shear correlation function statistic to the more stringent `COSEBI' statistic \citep{schneider/etal:2010} renders the cosmic shear measurement completely insensitive to the low-$\ell$ scales where these approximations become important.

>>>>

In this comment we have highlighted a number of shortcomings in the analysis presented in \cite{kitching/etal:2016}.  We argue that the flat sky and Limber approximations that can be used in cosmic shear analyses are not able to explain away the tension between current weak lensing and Planck results, in contrast to the message that \cite{kitching/etal:2016} convey.  Our critique should not, however, detract from the important core message of their paper, that future surveys will need to consider these approximations and optimise their statistical analyses accordingly (\shahab{I think we are being too nice, it's unclear the Limber and flat-sky approximations are important even for future surveys when considered together --- moreover Tommaso's paper seems to suggest Limber approximation fine even for future surveys}).  For example moving from the standard two-point shear correlation function statistic to the more stringent `COSEBI' statistic \citep{schneider/etal:2010} renders the cosmic shear measurement completely insensitive to the low-$\ell$ scales where these approximations become important.
